---
title: "Public Schools Project - Score-Frequency Graphics"
author: "D. Hopp"
date: "September 2017"
output:
  html_document: 
    css: "my_css.css"
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width=100)
```  

```{r, echo=FALSE, warning=FALSE, message=FALSE}
require(rprojroot)
require(reshape2)
require(magrittr)
require(plyr)
require(tidyverse)
require(cowplot)
require(grid)
require(readr)
require(taRifx)
require(R.utils)
require(ggthemes)
require(emdist)
require(rlist)
require(MASS)
require(testthat)
library(printr)
```  

[comment]: # (use fn_pltScoreFreqNCDPI)

```{r fetch, echo=FALSE, message=FALSE}
load("3.RData")
```  

```{r grab_grobs, echo=FALSE}
plts_ <- fn_pltScoreFreqNCDPI(argCOW=TRUE,argSave=FALSE)
```  

#### Each plot shows the five achievement levels, reading from left (Level 1) to right (Level 5). Scoring in Level 3, 4, or 5 is Grade Level Proficient.

The changes seen in score frequencies are due either to differences in test difficulties, or differences in the abilities of students, or some combination of both. Random effects play a role, since the same students, taking the same tests, on some other day would be expected to exhibit some differences in their scores.  

```{r plot_this_1, echo=FALSE, fig.width=10}
cowplot::plot_grid( plts_[[1]], plts_[[7]], plts_[[13]], ncol=3)
```  

 ```{r plot_this_2, echo=FALSE, fig.width=10}
cowplot::plot_grid( plts_[[2]], plts_[[8]], plts_[[14]], ncol=3)
```  

 ```{r plot_this_3, echo=FALSE, fig.width=10}
cowplot::plot_grid( plts_[[3]], plts_[[9]], plts_[[15]], ncol=3)
```  

```{r plot_this_4, echo=FALSE, fig.width=10}
cowplot::plot_grid( plts_[[4]], plts_[[10]], plts_[[16]], ncol=3)
```  

 ```{r plot_this_5, echo=FALSE, fig.width=10}
cowplot::plot_grid( plts_[[5]], plts_[[11]], plts_[[17]], ncol=3)
```  

 ```{r plot_this_6, echo=FALSE, fig.width=10}
cowplot::plot_grid( plts_[[6]], plts_[[12]], plts_[[18]], ncol=3)
```  

### Cut Points  
The cut points are consistent across the three years.  
p50 is the weighted median. It is very stable, and is located in Level 3, that is, where students classify as Grade Level Proficient, but not Career and College Ready.  

#### For ELA  
```{r cut_points_1, echo=FALSE}
dplyr::inner_join(
  dfScoreFreqNCDPIBreaks,
  dfScoreFreqNCDPIquartiles %>%
    dplyr::select(
      acYear,
      Grade,
      Subject,
      p50
    ),
  by=c(
    "acYear",
    "Grade",
    "Subject")
) %>%
  dplyr::filter(Subject=="ELA" & Grade=="3") %>%
  dplyr::arrange(f_acYear) %>%
  dplyr::select(-f_acYear)
```  

```{r cut_points_2, echo=FALSE}
dplyr::inner_join(
  dfScoreFreqNCDPIBreaks,
  dfScoreFreqNCDPIquartiles %>%
    dplyr::select(
      acYear,
      Grade,
      Subject,
      p50
    ),
  by=c(
    "acYear",
    "Grade",
    "Subject")
) %>%
  dplyr::filter(Subject=="ELA" & Grade=="4") %>%
  dplyr::arrange(f_acYear) %>%
  dplyr::select(-f_acYear)
```  

```{r cut_points_3, echo=FALSE}
dplyr::inner_join(
  dfScoreFreqNCDPIBreaks,
  dfScoreFreqNCDPIquartiles %>%
    dplyr::select(
      acYear,
      Grade,
      Subject,
      p50
    ),
  by=c(
    "acYear",
    "Grade",
    "Subject")
) %>%
  dplyr::filter(Subject=="ELA" & Grade=="5") %>%
  dplyr::arrange(f_acYear) %>%
  dplyr::select(-f_acYear)
```  

#### For MA  
```{r cut_points_4, echo=FALSE}
dplyr::inner_join(
  dfScoreFreqNCDPIBreaks,
  dfScoreFreqNCDPIquartiles %>%
    dplyr::select(
      acYear,
      Grade,
      Subject,
      p50
    ),
  by=c(
    "acYear",
    "Grade",
    "Subject")
) %>%
  dplyr::filter(Subject=="MA" & Grade=="3") %>%
  dplyr::arrange(f_acYear) %>%
  dplyr::select(-f_acYear)
```  

```{r cut_points_5, echo=FALSE}
dplyr::inner_join(
  dfScoreFreqNCDPIBreaks,
  dfScoreFreqNCDPIquartiles %>%
    dplyr::select(
      acYear,
      Grade,
      Subject,
      p50
    ),
  by=c(
    "acYear",
    "Grade",
    "Subject")
) %>%
  dplyr::filter(Subject=="MA" & Grade=="4") %>%
  dplyr::arrange(f_acYear) %>%
  dplyr::select(-f_acYear)
```  

```{r cut_points_6, echo=FALSE}
dplyr::inner_join(
  dfScoreFreqNCDPIBreaks,
  dfScoreFreqNCDPIquartiles %>%
    dplyr::select(
      acYear,
      Grade,
      Subject,
      p50
    ),
  by=c(
    "acYear",
    "Grade",
    "Subject")
) %>%
  dplyr::filter(Subject=="MA" & Grade=="5") %>%
  dplyr::arrange(f_acYear) %>%
  dplyr::select(-f_acYear)
```  
